Geometric models reveal the hidden structure of conceptual knowledge

### 2023-01-25 paper/main.tex
added paragraph with missing stats for knowledge timeseries figure and associated text

The ability to quantify how much each question is ``asking about'' each moment
of the lectures could enable high-resolution insights into participants'
knowledge. Traditional approaches to estimating how much a student ``knows''
about the content of a given lecture entail computing the proportion of
correctly answered questions. But if two students receive identical scores on
an exam, might our modeling framework help us to gain more naunced insights
into the \textit{specific} content that each student has mastered (or failed to
master)? For example, a student who misses three questions that were all about
the same concept (e.g., concept $A$) will have gotten the same
\textit{proportion} of questions correct as another student who missed three
questions about three \textit{different} concepts (e.g., $A$, $B$, and $C$).
But if we wanted to fill in the ``gaps'' in the two students' understandings,
we might do well to focus on concept $A$ for the first student, but add in
concepts $B$ and $C$ for the second student.

\textbf{JRM STOPPED HERE...}

As a sanity check, we anticipated that our knowledge estimates
should show a content-specific ``boost'' in participants' knowledge after
watching each video.

@@

@@

    \caption{\textbf{Estimating knowledge for held-out questions.}}

>>>>

The ability to quantify how much each question is ``asking about'' each moment
of the lectures could enable high-resolution insights into participants'
knowledge. Traditional approaches to estimating how much a student ``knows''
about the content of a given lecture entail computing the proportion of
correctly answered questions. But if two students receive identical scores on
an exam, might our modeling framework help us to gain more naunced insights
into the \textit{specific} content that each student has mastered (or failed to
master)? For example, a student who misses three questions that were all about
the same concept (e.g., concept $A$) will have gotten the same
\textit{proportion} of questions correct as another student who missed three
questions about three \textit{different} concepts (e.g., $A$, $B$, and $C$).
But if we wanted to fill in the ``gaps'' in the two students' understandings,
we might do well to focus on concept $A$ for the first student, but to also add
in materials pertaining to concepts $B$ and $C$ for the second student.

We developed a simple formula (Eqn.~\ref{eqn:prop}) for using a participant's
responses to a small set of multiple choice questions to estimate how much the
participant ``knows'' about the concept reflected by any arbitrary coordinate,
$x$, in text embedding space (e.g., the content reflected by any moment in a
lecture they had watched; see \textit{Estimating dynamic knowledge traces}).
Essentially, the estimated knowledge at the coordinate is given by the weighted
average proportion of quiz questions the participant answered correctly, where
the weights reflect how much each question is ``about'' the content at $x$.
When we apply this approach to estimate the participant's knowledge about the
content presented in each moment of each lecture, we can obtain a detailed
timecourse describing how much ``knowledge'' the participant has about any part
of the lecture. As shown in Figures~\ref{fig:knowledge-timeseries}A and C), we
can also apply this approach separately for the questions from each quiz the
participants took throughout the experiment. From just 13 questions per quiz,
we obtain a high-resolution snapshot (at the time each quiz was taken) of what
the participants knew about any moment's content, from either of the two
lectures they watched (comprising a total of 1106 samples across the two
lectures).

@@

Of course, even though the timecourses in
Figure~\ref{fig:knowledge-timeseries}A and C provide detailed
\textit{estimates} about participants' knowlege, those estimates are only
\textit{useful} to the extent that they accurately reflect what participants
actually know. As one sanity check, we anticipated that the knowledge estimates
should show a content-specific ``boost'' in participants' knowledge after
watching each lecture. In other words, if participants learn about each
lecture's content when they watch each lecture, the knowledge estimates should
reflect that. After watching the \textit{Four Fundamental Forces} lecture,
participants should show more knowledge for the content of that lecture than
they had before, and that knowledge should persist for the remainder of the
experiment. Specifically, knowledge about that lecture's content should be
relatively low when estimated using Quiz 1 responses, but should increase when
estimated using Quiz 2 or 3 responses (Fig.~\ref{fig:knowledge-timeseries}B).
Indeed, we found that participants' estimated knowledge about the content of
the \textit{Four Fundamental Forces} was substantially higher on Quiz 2 versus
Quiz 1 ($t(XX) = XX, p = XX$) and on Quiz 3 versus Quiz 1 ($t(XX) = XX, p =
XX$). We found no reliable differences in estimated knowledge about that
lecture's content on Quiz 2 versus 3 ($t(XX) = XX, p = XX$). Similarly, we
hypothesized (and subsequently confirmed) that participants should show more
estimated knowledge about the content of the \textit{Birth of Stars} lecture
after (versus before) watching it (Fig.~\ref{fig:knowledge-timeseries}D).
Specifically, since participants watched that lecture after taking Quiz 2 (but
before Quiz 3), we hypothesized that their knowledge estimates should be
relatively low on Quizzes 1 and 2, but should show a ``boost'' on Quiz 3.
Consistent with this prediction, we found no reliable differences in estimated
knowledge about the \textit{Birth of Stars} lecture content on Quizzes 1 versus
2 ($t(XX) = XX, p = XX$), but the estimated knowledge was substantially higher
on Quiz 3 versus 2 ($t(XX) = XX, p = XX$) and Quiz 3 versus 1 ($t(XX) = XX, p =
XX$).

@@

    \caption{\textbf{Estimating knowledge at the embedding coordinates of
    held-out questions.} \textbf{Top panels.} Separately for each quiz (panel
    column), we plot the average correlations between the topic vectors of
    held-out correctly (blue) and incorrectly (red) answered questions, and the
    topic vectors for the $n$ closest remaining correctly or incorrectly
    answered questions, respectively. For details, see \textit{Testing the
    ``smoothness'' of knowledge}. \textbf{Bottom panels.} Separately for each
    quiz (panel column), we plot the distributions of predicted knowledge at
    the embedding coordinates of each held out correctly (blue) or incorrectly
    (red) answered question. The $t$-tests reported in each panel are between
    the distributions of estimated knowledge at the coordinates of correctly
    versus incorrectly answered held-out questions. For details, see
    \textit{Estimating held-out conceptual knowledge}.}
